Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-14 cover_letter/letterhead.tex
cover letter and citations

\long\def\closing#1{
\vspace{0.1in} % Some whitespace after the letter content and before the signature
\noindent % Stop paragraph indentation
%\hspace*{\longindentation} % Move the signature right
%\parbox{\indentedwidth}{\raggedright
#1 % Print the signature text
\vskip 0.0in
\includegraphics[width=0.25\textwidth]{signature_720}
%\vskip 0.65in % Whitespace between the signature text and your name
\\
\noindent
\fromsig
\vskip -0.05in
\titleone\\
\department\\
\institution\\
\vskip -0.1in
\noindent cc: M.~Zhao, K.~Lachowski, S. Alamdari, J.~Sampath, P.~Mu, L.D.~Pozzo, C.-L.~Chen, C.J.~Mundy, J.~Pfaendtner 
}

@@

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

@@

We are pleased to submit a manuscript for consideration for publication in the \textit{The Journal of the American Chemical Society} an original research \textbf{Article} entitled \textbf{``Hierarchical Self-Assembly Pathways of Polypeptoid Helices and Sheets''} by Mingfei Zhao (U.~Chicago), Kacper Lachowski(U.W.~Seattle), Sarah Alamdari (U.W.~Seattle), Janani Sampath (PNNL),  Peng Mu (PNNL), Chun-Long Chen (PNNL, U.W.~Seattle), Lilo D.\ Pozzo (U.W.~Seattle), Christopher J.\ Mundy (PNNL, U.W.~Seattle), Jim Pfaendtner (U.W.~Seattle, PNNL), and Andrew L.\ Ferguson (U.~Chicago). This article has not been published previously and is not under consideration in any other journal. All authors have seen and approved the submission of the manuscript.

% background and why it is important to do this.
\textbf{Background.} Polypeptoids (poly-N-substituted glycines) are a class of highly tailorable synthetic peptidomic polymers with applications as drugs, antimicrobials, and catalysts that can be engineered to assemble into nanoaggregates including spheres, helices, tubes, and sheets. Amphiphilic diblock polypeptoids have been engineered to assemble high-aspect ratio 2D crystalline lattices with applications in catalysis, molecular separations, and photovoltaics. Assembly is induced by dissolving the peptoids in an organic solvent/water mixture and evaporating the organic phase to promote hydrophobic association, but the mechanistic pathways and thermodynamic/kinetic driving forces mediating assembly remain uncharacterized.

% what we did in this paper
\textbf{Key Results.}  In this work, we integrate all-atom molecular dynamics (MD) simulation, small angle x-ray scattering (SAXS), circular dichroism (CD), atomic force microscopy (AFM), and x-ray diffraction (XRD) to present an integrated computational/experimental study of the self-assembly of Nbrpe6Nc6 -- an amphiphilic diblock polypeptoid comprising an \ce{NH2} capped block of six hydrophobic N-((4-bromophenyl)ethyl)glycine residues conjugated to a hydrophilic 6-aminohexanal (\ce{NH3(CH2)5CO}) tail -- in organic solvent/water mixtures. Our MD simulations predict a hierarchical assembly pathway by which monomers first assemble into disordered aggregates that self-order into 1D chiral helical rods that subsequently assemble 2D achiral crystalline nanosheets. Experimentally, we observe stable, free-floating 1D rods in mixed solvent using SAXS and CD, and 2D crystalline sheets in pure water using XRD and AFM. MD predictions of the relative thermodynamic stability of disordered aggregates, helical rods, and sheets as a function of the organic solvent concentration support a thermodynamically-controlled assembly process.

% impact and importance of this paper
\textbf{Significance.} Our work integrates simulation and experiment to establish a previously unknown hierarchical polypeptoid assembly pathway wherein 0D monomers first assemble into disordered aggregates that ripen into 1D helical rods that subsequently form 2D crystalline nanosheets. This new understanding of the hierarchical assembly mechanisms and stable intermediaries presents new principles and guidance for the rational design of peptoid-based nanomaterials with potential biochemical, biomedical, and bioengineering applications including as drugs, antimicrobials, catalysts, and molecular sieves.

% fit for the journal
\textbf{Scope.}  This work fits within the scope of \textit{The Journal of the American Chemical Society} in that it reports new fundamental physical chemical understanding of the molecular pathways and mechanisms of polypeptoid self-assembly. This new understanding establishes new mechanistic understanding and a thermodynamic basis for existing experimental evaporation-induced assembly protocols and identifies previously unknown solvent conditions for the stable fabrication of 1D helical rods. These principles offer new strategies for the supramolecular assembly of diverse peptoid sequences into 1D and 2D supramolecular aggregates.

@@

\textbf{\textit{Simulation.}}

Stephen Whitelam \hfill Molecular Foundry, Lawrence Berkeley National Lab \\
(510) 495-2769 \hfill \url{swhitelam@lbl.gov}

$\circ$ \textit{Expert in biomolecular simulation and molecular modeling of peptoid assembly.}

Monica Olvera de la Cruz \hfill Materials Science and Engineering and Chemistry, Northwestern University \\
(847) 491-7801 \hfill \url{m-olvera@northwestern.edu}

$\circ$ \textit{Expert in molecular modeling and simulation of peptides, polymers, and materials assembly.}

Arthi Jayaraman \hfill Chemical and Biomolecular Engineering, University of Delaware \\
(302) 831-8682 \hfill \url{arthij@udel.edu}

$\circ$ \textit{Expert in theory and simulation of biomolecular assembly and molecular materials design.}

@@

Darrin Pochan \hfill Materials Science and Engineering, University of Delaware \\
(302) 831-3569 \hfill \url{pochan@udel.edu}

$\circ$ \textit{Expert in biomolecular and polymeric assembly and characterization.}

Rachel Segalman \hfill Chemical Engineering, UC Santa Barbara \\
(805) 893-3709 \hfill \url{segalman@ucsb.edu}

$\circ$ \textit{Expert in biological and polymeric materials assembly.}

Ron Zuckermann \hfill The Molecular Foundry, Lawrence Berkeley National Laboratory \\
(510) 387-9364 \hfill \url{rnzuckermann@lbl.gov}

$\circ$ \textit{Pioneer of peptoid science. Expert in peptoid synthesis and characterization.}

>>>>

\long\def\closing#1{
\vspace{0.1in} % Some whitespace after the letter content and before the signature
\noindent % Stop paragraph indentation
%\hspace*{\longindentation} % Move the signature right
%\parbox{\indentedwidth}{\raggedright
#1 % Print the signature text
\vskip 0.0in
\includegraphics[width=0.25\textwidth]{signature_720}
%\vskip 0.65in % Whitespace between the signature text and your name
\\
\noindent
\fromsig
\vskip -0.05in
\titleone\\
\department\\
\institution\\
\vskip -0.1in
\noindent cc: M.~Jones, B.~Ashwood, A.~Tokmakoff
}

@@

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

We are pleased to submit a manuscript for consideration for publication in the \textit{The Journal of the American Chemical Society} an original research \textbf{Article} entitled \textbf{``Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy''} by Michael S. Jones (U.~Chicago), Brennan Ashwood(U.~Chicago), Andrei Tokmakoff(U.~Chicago), and Andrew L.\ Ferguson (U.~Chicago). This article has not been published previously and is not under consideration in any other journal. All authors have seen and approved the submission of the manuscript.

% background and why it is important to do this.
\textbf{Background.}  A robust understanding of the sequence-dependent thermodynamics of DNA hybridization has enabled rapid advances in DNA nanotechnology. A fundamental understanding of the sequence-dependent kinetics and mechanisms of hybridization and dehybridization remains comparatively underdeveloped. Although previous experimental and computational studies have identified deviations from an "all-or-nothing" hybridization/dehybridization model, few studies have integrated these approaches in order to evaluate and quantify sequence-dependent dynamics.

% what we did in this paper
\textbf{Key Results.}  In this work, we establish new understanding of the sequence-dependent hybridization/dehybridization kinetics and mechanism within a family of self-complementary pairs of 10-mer DNA oligomers by integrating coarse-grained molecular simulation, machine learning of the slow dynamical pathways, data-driven inference of long-time kinetic models, and experimental temperature-jump infrared spectroscopy. We conducted more than 1 ms of unbiased coarse-grained molecular dynamics simulations and employed deep learning techniques to construct high-resolution Markov state models as predictive and interpretable models of the sequence dependent dynamics. Our results reveal that the specific placement of interrupting G:C pairs within an otherwise repetitive AT sequence can have a profound impact on the kinetic pathways and mechanisms for association and dissociation of the DNA duplex.

% impact and importance of this paper
\textbf{Significance.} Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation in burgeoning DNA nanotechnology applications.

% fit for the journal
\textbf{Scope.}  This work fits within the scope of \textit{The Journal of the American Chemical Society} in that it reports new fundamental physical chemical understanding of the molecular pathways and mechanisms of DNA hybridization as a function of sequence. These sequence-dependent mechanisms are of increasing importance to both fundamental biophysical and biochemical processes and rapidly developing nanotechnology applications.

@@

\textbf{\textit{Simulation.}} \noteb{Notes are just to show connection}

Joshua Lequieu \hfill Chemical and Biological Engineering, Drexel University \\
(215) 895-6624 \hfill \url{lequieu@drexel.edu}

$\circ$ \textit{Expert in multi-scale biomolecular simulation with a focus in materials science and biology. \noteb{3spn2}}

Jonathon K. Whitmer \hfill Chemical and Biomolecular Engineering, Notre Dame \\
(574) 631-1417 \hfill \url{jwhitme1@nd.edu}

$\circ$ \textit{Expert in simulations of polyelectrolytes and self-assembling systems.  \noteb{3spn2}}

Thomas Knotts \hfill Chemical Engineering, Brigham Young University \\
(801) 422-9158 \hfill \url{thomas.knotts@byu.edu}

$\circ$ \textit{Expert in theory and simulation of multi-molecular systems.  \noteb{3spn.0 and .1}}

Shoji Takada \hfill Biophysics, Graduate School of Science, Kyoto University \\
+81-(0)75-753-4070 \hfill \url{takada@ biophys.kyoto-u.ac.jp}

$\circ$ \textit{Expert in simulations of DNA-protein systems \noteb{3spn2 and  CG protein modeling}}

Thomas Ouldridge \hfill  Bioengineering, Imperial College London \\
+44-(0)20-7589-5111 \hfill \url{t.ouldridge@implerial.ac.uk}

$\circ$ \textit{Expert in simulations of biochemical systems for nanotechnology applications \noteb{oxDNA}}

Modesto Orozco \hfill  Institute for Research in Biomedicine, University of Barcelona \\
+34-93-40-37156 \hfill \url{modesto.orozco@irbbarcelona.org}

$\circ$ \textit{Expert in molecular dynamics simulations of nucleic acids \noteb{all-atom DNA model}}

@@

Anjum Ansari \hfill Physics, University of Illinois at Chicago\\
(312) 996-8735 \hfill \url{ansari@uic.edu}

$\circ$ \textit{Expert in kinetics of biomolecular relaxation from non-equilibrium ensembles.}

Neil Hunt \hfill Chemistry, University of York \\
+44-(0)19-0432-2511 \hfill \url{neil.hunt@york.ac.uk}

$\circ$ \textit{Expert in ultrafast spectroscopy of complex fluids and biological systems.}

Xinsheng Zhao \hfill Chemistry and Molecular Engineering, Peking University \\
+86-10-6275-1727 \hfill \url{zhaoxs@pku.edu.cn}

$\circ$ \textit{Expert in dynamic mechanisms of biomolecular reactions and nucleic acid biophysics.}
